CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility

### 2021-02-01 codecheck-paper.tex
integrate editorial review by CB
also adjust wording in figure

\begin{document}
%\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles}
\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility}
\author[1,$\ast$]{Daniel~N\"{u}st \orcid{0000-0002-0024-5046} \github{nuest} \twitter{nordholmen}}
\author[2,$\ast$]{Stephen~J~Eglen \orcid{0000-0001-8607-8025} \github{sje30} \twitter{StephenEglen}}
\affil[1]{Institute for Geoinformatics, University of M\"unster, Germany; \email{daniel.nuest@uni-muenster.de}}
\affil[2]{Department of Applied Mathematics and Theoretical Physics, University of Cambridge, GB; \email{sje30@cam.ac.uk}}
\affil[$\ast$]{Contributed equally}
\affil[$$]{Public project for this article is at \url{https://github.com/codecheckers/paper}.}
\maketitle
\begin{abstract}
  The scientific paper falls short of effectively communicating computational research.
  To break out of the stalemate between multiple lacks of incentives, guidance, benefits, recognitions, and reproducibility, we propose to introduce execution of workflows underlying research articles as part of peer review.
  The CODECHECK system relies on five principles and can be integrated into review and publications processes in multiple ways, which we describe along multiple dimensions (importance, who, openness, when).
  We demonstrate CODECHECK with 25 reproductions of diverse scientific publications in collaboration with academic publishers and conferences.
  These CODECHECKs show that asking for reproducible workflows in a collaborative review can effectively improve executability.
  CODECHECK relies on free infrastructure and open tools.
  While it has clear limitations, it can be a building block in the open-science and publishing ecosystems to improve the reproducibility, appreciation, and potentially quality of non-textual research artefacts.
\end{abstract}

@@

Many areas of scientific research use computations to either simulate
or analyse their data. These computations are increasingly complex, and
difficult to explain coherently in a paper \citep{marwick_how_2015}.
To complement the traditional route of sharing research by writing papers,
there is a growing demand to share the underlying artefacts, notably 
code and datasets, so that others can inspect, reproduce or expand that work
(see Figure~\ref{fig:inverse}).
Some of the earliest proponents were Buckheit and Donoho
\cite{buckheit_wavelab_1995} who coined what has been called 
\emph{Claerbout's claim} (extending on \citet{claerbout_electronic_1992}):
\emph{"An article about computational science in a scientific publication 
is \text{not} the scholarship itself, it is merely \textbf{advertising} of
the scholarship. The actual scholarship is the complete software development
environment and the complete set of instructions which generated the 
figures."}
%  Donoho (2010) has another paraphrasing of the same quote, which might work better:
% "an article about computational result is advertising, not scholarship. The actual scholarship is the full software environment, code and data, that produced the result."

So, assuming that researchers begin to share more artefacts, how might
they be examined to test that they do what they claim? For example,
most scientific journals now require a data sharing statement that
outlines what data the authors have (or will) share.
The implementation status of this varies considerably according to the
journal. At one end of the spectrum, there are specialist journals that
have been created to accept ``data papers''
(e.g.~\href{https://www.nature.com/sdata/}{\emph{Scientific~Data}}, 
\href{https://essd.copernicus.org/}{\emph{Earth System Science Data}}, 
\href{https://rmets.onlinelibrary.wiley.com/journal/20496060}{\emph{Geoscience Data Journal}},
\href{https://bdj.pensoft.net/}{\emph{Biodiversity Data Journal}},
\href{https://openpsychologydata.metajnl.com/}{\emph{Journal of Open Psychology Data}},
\href{https://odjar.org/}{\emph{Open Data Journal for Agricultural Research}},
\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}}).
These journals have established rigorous procedures by which
the data is validated according to standards in each field. At the
other end of the spectrum, we still witness today the infamous
statement ``Data available upon reasonable request''. Authors,
while possible well-intentioned at the time of writing the article,
often cannot provide the data when readers ask for
it as data disappears over time \cite{Vines2014-hf}.

Given that there are no clear standards yet for sharing data, what hope
might there be for sharing computer programs? While both data and software
together are required to validate outcomes of a computational analysis,
they do differ in  that data can be seen as static and inert,
while code requires an environment to be run in and interacted with.
This makes software harder to share.
Our anectodal experiences
on this matter suggest that often a variety of reasons are given for why
code cannot be shared, e.g.,~``there is no documentation'', ``I cannot 
maintain it'', or ``I do not want to give away my code to competitors''.
Our view, rooted in Open~Science principles
\footnote{\url{https://en.wikipedia.org/wiki/Open_science}}, is that
wherever possible sharing code is good for the community and the 
individual \cite{Barnes2010-iv,markowetz_five_2015}.
Having code and data openly available and archived, 
provides a valuable resource for others to learn
from, even when it doesn't run anymore or if there is no documentation.
However,
with a little effort, we believe that if someone independent can execute
an analysis, this is worth documenting as early as possible and that the barrier
to evaluating non-text materials of research can be greatly reduced with a 
focused approach.
Just like data journals' validations of data and all journals' peer review
can provide a ``baseline reassurance'', i.e.,~a paper
has been checked by someone with an understanding of the topic
\cite{fyfe_mission_2019}, we think the same baseline should be provided
regarding the workflow underlying a paper.
With this in mind, we have developed a set of principles and an example 
workflow that provides what he hope is a pragmatic way of checking that a
paper's code works.

The main contribution of this work is a thorough description of a process
and its variations to integrate a much needed evaluation of the 
computational reproducibility\footnote{Following the 
Claerbout/Donoho/Peng terminilogy, see 
\cite{barba_terminologies_2018}} into peer
review, and a demonstration of its feasibility by means of 25
reproductions across different scientific disciplines.
We call this system CODECHECK as outlined next.

\begin{figure}
% emojis: https://github.com/googlefonts/noto-emoji
% cross mark: u274c
% check mark: u2714
  \centering
  \includegraphics[width=0.7\textwidth]{figs/rr.pdf}
  \caption{The inverse problem in reproducible research.  The left
  half of the diagram shows the diverse range of materials used
  within a laboratory.  These materials are often then
  condensed for sharing with the outside world via the
  research paper, a static PDF document.  Working backwards from the
  PDF to the underlying materials is impossible. This prohibits reuse
  and is not only non-transparent for a specific paper, but also 
  ineffective for science as a whole. By sharing the
  materials on the left, others outside the lab can reproduce
  or build on this work.}
  \label{fig:inverse}
\end{figure}

@@

CODECHECK is best demonstrated by way of our example workflow, and later
we expand on the underlying principles. The workflow involves three
groups of people:
(1) the AUTHOR of a paper providing the code to be checked,
(2) the PUBLISHER of a journal interested in publishing the paper 
by the author, and
The workflow we have refined is shown in Figure~\ref{fig:workflow}
and consists of six steps.

@@

\textbf{Step 1:} the author submits their manuscript along with code and data to
the PUBLISHER. The code and data need not be openly available at this point.
However, in many cases the code and data may be published on a code hosting platform,
such as GitHub or GitLab. Ideally, the AUTHOR is expecting the CODECHECK and
prepares for it, e.g.,~by asking a colleague to attempt a reproduction or by 
following the PUBLISHER's guidelines for sharing computational workflows.

\textbf{Step 2:} the publisher finds a CODECHECKER to check the code. This is
analogous to the publisher finding one or more reviewers to evaluate
the paper. However, a crucial difference is that we suggest the CODECHECKER
and the author can talk freely to each other directly (see next), independent
of the openness of the scientific peer review process.

\textbf{Step 3:} the CODECHECKER runs the code, based on instructions provided by
the author. They check if some or all of the results from the paper can be
reproduced. If there are
any problems running the code, the CODECHECKER asks the AUTHOR for help,
updates, or further documentation.
The burden to provide reproducible material lies with the author.
The CODECHECKER then tries to run the code again.
This process iterates until either the CODECHECKER is successful,
or the CODECHECKER concludes the code does not work.
As part of this process, the CODECHECKER could work entirely
locally on their own compute resource, or in the cloud, e.g.,~using the open
MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which
are more tailored for scientific publications while others offer commercial
options to, e.g.,~publishers (cf. \cite{konkol_publishing_2020}).
Cloud-based 
infrastructure allows for the CODECHECKER and author to collaboratively improve
the code and enforces a complete definition of the computing environment,
but, unless secure infrastructure is provided, e.g.,~by the PUBLISHER,
requires code and data to be published openly online.

Note that the task of the CODECHECKER is to check the ``mechanics'' of the
workflow, not to evaluate its correctness. 
Stodden et~al. \cite{stodden_setting_2013} distinguish between \emph{verification}
and \emph{validation}
in the context of Mathematics. Following their definition, a CODECHECK
ensures a verification\footnote{Note this is a different use as the
\emph{Verification Reports} 
article type in the journal \emph{Cortex} \cite{chambers_verification_2020}},
of computational results, i.e.,~checking that the code correctly produces the 
output it claims to create, but not a validation, i.e.,~does the code implement
the right algorithm to solve the problem that is researched.
Nevertheless, the simple attempt to reproduce may highlight shortcomings in
a submission to meet journal policy requirements
(cf.~\cite{christian_journal_2020})
and can effectively increase transparency and thereby improve practices
(cf.~\cite{nosek_scientific_2012})
even if the check goes not into all details.

\textbf{Step 4:} the CODECHECKER writes a certificate stating how the code was
run and includes a copy of outputs (figures, tables) that were
independently generated. The certificate may include further recommendations to
the author how to improve the material.
The free text in the certificate is the most flexible and feasible way to 
make clear what was really checked, because each workflow is unique.
Since no specific tool or platform should be required so that no authors are
excluded, automation or fixed checklists are futile.

\textbf{Step 5:} the certificate as well as auxiliary files created
during the check, e.g.,~a specification of a computing environment, data 
subsets or helper scripts, and the original code and data get deposited in
an open archive unless restrictions regarding data size, license or 
sensitivity apply.
Currently, the CODECHECKERs deposit the material on Zenodo themselves, 
while processes managed by the PUBLISHER may have them complete this step.
A badge or other visual aid may be added to the deposit and the paper and
link to the certificate.
Despite the simplification into a binary value and the risk of confusion of
the extent of the check, especially across disciplines and methods,
a badge can provide recognition value and creditation for the completed
CODECHECK. The badge and the actual check are incentives for the efforts
of providing a readily computationally reproducible workflows.

\textbf{Step 6:} the PUBLISHER can, depending on the timing, provide the
certificate to scientific reviewers or handling editors, or publish it
and link between the certificate, the paper, and any repositories 
through metadata. Currently, the CODECHECKER creates these connections on
Zenodo \footnote{
See ``related identifiers'' and ``alternate identifiers'' in the right-hand
sidebar of certificate \texttt{2020-025} \cite{cert-2020-025}.}.
The PUBLISHER also gives credit to the CODECHECKER for their work by depositing
the activity appropriately in scholarly profiles, such as ORCID\footnote{
See peer review contributions in ORCID records at
\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}.}.
The PUBLISHER also ensures proper publication metadata, e.g.,~links from the 
certificate repository to the published paper or the original code repository.

@@

The example workflow leaves room for different
implementations of this process. To elaborate on these options, we consider
the several dimensions in a space of possible CODECHECK workflows, as shown in 
Figure~\ref{fig:dimensions}, to represent relevant aspects for the involved
stakeholders. These aspects touch on timing, responsibilities, and 
transparency and will be discussed in the following sections.
The options on each dimension have their own pros and cons and can,
for the most part, be freely combined.

@@

The point in time for a CODECHECK and the ascribed importance are closely
connected, so the dimensions \emph{When} and \emph{Imporance} are described
together.
The earlier a CODECHECK happens in the process, the more the
results can play a role in editorial decisions:
is a paper published, sent back for revisions, or rejected.
Even earlier checks, i.e.,~a CODECHECK of a preprint, may even help to 
improve the workflow itself even before a publisher is involved. The 
possibility for codechecking papers could be part of a preprint server's
policy, or simply be initiated by interested readers and community members.

A \textbf{strict prerequisite} for passing a check can reduce the workload
of reviewers, and therefore seems out of place late in the review process.
The later in a review process the check happens, the less challenging is it
to allow a bidirectional communication between author and codechecker, e.g.,
because the author might already be notified of the acceptance and may more
freely share materials online closer to the paper publication date.

A \textbf{pre-review} CODECHECK allows editors to use it as a filter to only
send  a submission out to peer review if it passes the check.
The certificate would  be included in the submission package provided
to the reviewers, so they can include the CODECHECK's results in their 
evaluation of the work.
Reviewers may then judge the the relevance of the computations for the results
of the work, or follow journal guidelines as how strict to apply the check's
outcomes.

A CODECHECK may also be conducted in \textbf{parallel} to the scientific review.
This puts less burden on the turnaround time for the check, yet only makes the 
outcomes available for the final consideration of the handling editor.
The check could also be tasked after suggestion by a reviewer and remove the
need for a pre-review screening for computational workflows. However, 
soliciting such a ``specialist review'' is much less desirable than having
a regular CODECHECK, without any special treatment for some submissions.
In both cases, a CODECHECK could then be awarded \textbf{equal weight} as 
the other reviews for the editor's decision.

A \textbf{post acceptance} CODECHECK has the smallest impact on editorial 
decisions and may simply provide an \textbf{extra merit} on top of the 
acceptance of the submission. 
This  is the least impactful solution while still evaluating all material
and properly acknowledging the check results from the paper, because the 
check can be completed  before the publication of the paper.
The GIScience group of checks (see below) falls into this
category: the AGILE conference highlights articles where the reproducibility
was successfully reviewed with a badge on the volume and article landing
pages. Similarly, the GIScience papers in collaborations with journals were 
checked while the authors worked on revisions.

A CODECHECK may also be conducted \textbf{post publication}, though it
would required to update the article and article metadata later to reference 
the check, otherwise readers would presumably not be aware of it. In general,
publishers are hesitant to make such a revision of published articles.
While this has the least impact on the current publishing practices and also
enables open discourse,
it also has the smallest effect on the original submission
and does not acknowledge the importance of reproducible workflows for
ensuring good scientific practice.

In general, the timing of a CODECHECK needs to be connected with 
established or typical scholarly communication process.
While more radical ideas to change the paradigms
how researchers can share their work much more granually exist\footnote{
For example Octopus, 
\url{https://science-octopus.org/about}, or Hypergraph
\url{https://www.libscie.org/hypergraph}.},
% mimosa is another platform idea, but no real prototype AFAICT:
% https://projects.invisionapp.com/share/E9Z3RKE7W3P#/screens/436978924
% https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
there is a value in an evolutionary approach and we do not suggest
to completely remove the paper as a point in time to share work just yet.
Enhancing existing processes with CODECHECKs allows communities to 
transition towards more open practices at their own pace and ensure
no one is left behind unfairly.
When integrating a CODECHECK in existing review and publication processes,
the \emph{turnaround time} is crucial. Depending on when and who conducts
the check, it may be done faster, or add considerably to the overall
time from submission to publication.
In our first checks, we found that a codecheck may generally take from 2-5
hours, with some outliers on the higher end. This time includes writing and
publishing the certificate, but excludes the time for actual computations.
These efforts are comparable to the time needed to peer review a submission,
which concurs with the efforts some volunteer codecheckers are willing to
spend.
However, the time includes a considerable share of communicating about the
process,  especially around who publishes which document when, so that proper
cross-referencing between paper and certificate is ensured via persistent
identifiers.
When integrated into a peer review system, this handling of documents should
become much more effective and reduce organisational overhead.

\subsubsection*{Openness, or: Who knows who?}\label{who-knows-who}

The topic of anonymity is broadly discussed, not the least in the push towards
open peer review as part of the Open Science movement 
(cf. \cite{ross-hellauer_guidelines_2019}).
Without taking a strong stance on this discussion, the general motivation 
behind CODECHECK
for higher transparency and reproducibility does indeed favour a more open 
review process.
However, anonymity can also serve to protect individuals 
\cite{tennant_limitations_2020}, e.g.~junior scientists, whereas it might 
be acceptable to require revealing of names from tenured researchers.
The possible negative effects of a non-anonymous review are reduced
if a CODECHECK is not relevant for the decision to accept or reject, but that
is of course not desirable. Instead, we argue that a CODECHECK is technical
process that should in any case be fixable and not a question of opinion or
a faulty approach.
It is technically possible for a nonsense workflow to receive
a CODECHECK certificate.
If passing a CODECHECK becomes mandatory, the full transparency may have to be 
revisited as the relations between authors and codecheckers fall under the
same social and acommunity challenges as non-anonymous peer review
(cf.~\cite{everythinghertz123}).
% previous sentenc inspired by https://osf.io/9cftx/

The technical nature of the check and the challenge of good enough 
documentation is why we see great benefits in a effective bidirectional
means of communication between author and codechecker. Instead of trying to
fix problems or guess the next step, the codechecker can ask the author to 
rework the docs  or the update code.
Of course, the codechecker may also provide a fix within the code instead
of lengthily describing a small change if it is quickly done.
Authors and readers can profit from codecheckers' experience and approach,
as the latter may create useful and instructive files, e.g.,~a machine-readable
computing environment specification, during the check.
While communication between author and codechecker may be facilitated in an 
anonymous way by the 
publisher, it most likely only helps to protect the identity of the 
codechecker, because, especially when of a certain size, not trivial, or 
if very specialised, it is close to impossible to fully anonymise a codebase.
Therefore, the most effective means with the most desirable variant for
the stakeholders is a non-anonymous and collaborative CODECHECK.
The contributions by the codechecker may even be integrated into
the code of the workflow and be acknowledged as code commits. This way, 
proper credit is given within the research software development community.
Nevertheless, depending on needs represented by other dimensions,
the aspect of `Who knows who?' may be the most flexible one.

>>>>

\begin{document}
%\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles}
\title{CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility}
\author[1,$\ast$]{Daniel~N\"{u}st \orcid{0000-0002-0024-5046} \github{nuest} \twitter{nordholmen}}
\author[2,$\ast$]{Stephen~J~Eglen \orcid{0000-0001-8607-8025} \github{sje30} \twitter{StephenEglen}}
\affil[1]{Institute for Geoinformatics, University of M\"unster, Germany; \email{daniel.nuest@uni-muenster.de}}
\affil[2]{Department of Applied Mathematics and Theoretical Physics, University of Cambridge, GB; \email{sje30@cam.ac.uk}}
\affil[$\ast$]{Contributed equally}
\affil[$$]{Public project for this article is at \url{https://github.com/codecheckers/paper}.}
\maketitle
\begin{abstract}
  In its traditional format, the scientific paper falls short of effectively communicating computational research.
  Since this issue is difficult to address due to lacking incentives, guidance, benefits, recognitions, and reproducibility, we come at this problem from a different angle:
  We propose to introduce, as part of peer review, a system by which the workflows underlying research articles are executed.
  The CODECHECK system, which uses free intrastructure and open tools, relies on five principles and can be integrated into review and publication processes in multiple ways.
  We describe these integrations along multiple dimensions (importance, who, openness, when).
  In collaboration with academic publishers and conferences, we demonstrate CODECHECK with 25 reproductions of diverse scientific publications.
  These CODECHECKs show that asking for reproducible workflows during a collaborative review can effectively improve executability.
  While CODECHECK has clear limitations, it may represent a building block in the Open Science and publishing ecosystems for improving the reproducibility, appreciation, and, potentially, the quality of non-textual research artefacts.
\end{abstract}

@@

Many areas of scientific research use computations to either simulate or analyse their data. These computations are becoming increasingly complex and are difficult to explain coherently in a paper \citep{marwick_how_2015}.
To complement the traditional route of sharing research by writing papers,
there is a growing demand to share the underlying artefacts, notably 
code and datasets, so that others can inspect, reproduce or expand that work
(see Figure~\ref{fig:inverse}).
Some of the earliest proponents of this initiative were Buckheit and Donoho \cite{buckheit_wavelab_1995}, who coined what has been called \emph{Claerbout's claim} (extending on \citet{claerbout_electronic_1992}):
\emph{``An article about computational science in a scientific publication 
is \text{not} the scholarship itself, it is merely \textbf{advertising} of
the scholarship. The actual scholarship is the complete software development
environment and the complete set of instructions which generated the 
figures.''}

So, assuming that researchers begin to share more artefacts, how might these artefacts be examined to ensure that they do what they claim?
For example,
most scientific journals now require a data sharing statement that
outlines what data the authors have (or will) share.
Yet, each journal can implement this requirement differently.
At one end of the spectrum, journals have been created to accept ``data papers''
(e.g.,~\href{https://www.nature.com/sdata/}{\emph{Scientific~Data}}, 
\href{https://essd.copernicus.org/}{\emph{Earth System Science Data}}, 
\href{https://rmets.onlinelibrary.wiley.com/journal/20496060}{\emph{Geoscience Data Journal}},
\href{https://bdj.pensoft.net/}{\emph{Biodiversity Data Journal}},
\href{https://openpsychologydata.metajnl.com/}{\emph{Journal of Open Psychology Data}},
\href{https://odjar.org/}{\emph{Open Data Journal for Agricultural Research}},
\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}});
these journals have established rigorous procedures by which
the data is validated according to standards in each field.
At the other end of the spectrum, many journals still simply allow authors to state ``Data available upon reasonable request''.
Authors, while possibly well intentioned at the time of writing the article, often cannot provide the data when readers ask for it, as data disappears over time \cite{Vines2014-hf}.

Given that no clear standards yet exist for sharing data, what hope might there be for sharing computer programs?
Both data and software together are required to validate outcomes of a computational analysis, but they differ in that data can be seen as static and inert,
while code requires an environment to be run in and interacted with.
This makes software harder to share.
Our anectodal experiences on this matter suggest that researchers often give a variety of reasons for why code cannot be shared, e.g.,~``there is no documentation'', ``I cannot maintain it'', or ``I do not want to give away my code to competitors''.
Our view, rooted in Open~Science principles\footnote{\url{https://en.wikipedia.org/wiki/Open_science}}, is that sharing code, wherever possible, is good for the community and the individual \cite{Barnes2010-iv,markowetz_five_2015}.
Having code and data openly available and archived provides a valuable resource for others to learn from, even if the code no longer runs or if there is no documentation.
However, with a little effort, we believe that if an independent person can execute an analysis, this is worth documenting as early as possible and that such a focused approach can greatly reduce the barrier to evaluating non-text research materials.
Just as data journals' validations of data and all journals' peer review can provide a ``baseline reassurance'', i.e.,~that a paper has been checked by someone with an understanding of the topic \cite{fyfe_mission_2019}, we think the same baseline should be provided for the workflow underlying a paper.
With this in mind, we have developed a set of principles and an example workflow that provides what we hope is a pragmatic way of checking that a paper's code works.

As this work's main contribution, we offer a thorough description of a process and its variations to integrate a much needed evaluation of computational reproducibility\footnote{Following the Claerbout/Donoho/Peng terminology, see \cite{barba_terminologies_2018}.} into peer review, and we demonstrate its feasibility by means of 25 reproductions across different scientific disciplines.
We call this system CODECHECK as outlined next.

\begin{figure}
% emojis: https://github.com/googlefonts/noto-emoji
% cross mark: u274c
% check mark: u2714
  \centering
  \includegraphics[width=0.7\textwidth]{figs/rr.pdf}
  \caption{The inverse problem in reproducible research. The left
  half of the diagram shows the diverse range of materials used
  within a laboratory. These materials are often then
  condensed for sharing with the outside world via the
  research paper, a static PDF document. Working backwards from the
  PDF to the underlying materials is impossible. This prohibits reuse
  and is not only non-transparent for a specific paper but is also 
  ineffective for science as a whole. By sharing the
  materials on the left, others outside the lab can reproduce
  or build on this work.}
  \label{fig:inverse}
\end{figure}

@@

CODECHECK is best demonstrated by way of our example workflow, and later
we expand on the underlying principles. The workflow involves three
groups of people:
(1) the AUTHOR of a paper providing the code to be checked,
(2) the PUBLISHER of a journal interested in publishing the author's paper, and
(3) the CODECHECKER, who checks that the authors's code works.
The workflow we have refined is shown in Figure~\ref{fig:workflow}
and consists of six steps.

@@

\textbf{Step 1:} The author submits their manuscript along with the code and data to the PUBLISHER.
The code and data need not be openly available at this point.
However, in many cases the code and data may be published on a code hosting platform,
such as GitHub or GitLab. Ideally, the AUTHOR is expecting the CODECHECK and
prepares for it, e.g.,~by asking a colleague to attempt a reproduction or by 
following the PUBLISHER's guidelines for sharing computational workflows.

\textbf{Step 2:} The publisher finds a CODECHECKER to check the code. This is
analogous to the publisher finding one or more reviewers to evaluate
the paper.
However, a crucial difference is that we suggest the CODECHECKER and the author are able to talk freely and directly to each, regardless how open the scientific peer review process is.

\textbf{Step 3:} The CODECHECKER runs the code, based on instructions provided by
the author. They check if some or all of the results from the paper can be
reproduced. If there are
any problems running the code, the CODECHECKER asks the AUTHOR for help,
updates, or further documentation.
The burden to provide reproducible material lies with the author.
The CODECHECKER then tries to run the code again.
This process iterates until either the CODECHECKER is successful,
or the CODECHECKER concludes the code can not be executed within the context of a CODECHECK.
As part of this process, the CODECHECKER could work entirely locally, relying on their own computing resources, or in the cloud, e.g.,~using the open MyBinder infrastructure \cite{jupyter_binder_2018} or alternatives, some of which are more tailored to scientific publications while others offer commercial options for, e.g.,~publishers (cf. \cite{konkol_publishing_2020}).
A cloud-based infrastructure allows for the CODECHECKER and author to collaboratively improve the code and enforces a complete definition of the computing environment; but, unless secure infrastructure is provided, e.g.,~by the PUBLISHER, this requires the code and data to be published openly online.
Note that the task of the CODECHECKER is to check the ``mechanics'' of the workflow not to evaluate its correctness. 
In the context of mathematics, Stodden et~al. \cite{stodden_setting_2013} distinguish between \emph{verification} and \emph{validation};
following their definition, a CODECHECK ensures verification
\footnote{Note this is a different use from that referred to as the \emph{Verification Reports} article type in the journal \emph{Cortex} \cite{chambers_verification_2020}.},
of computational results, i.e.,~checking that the code correctly produces the output it claims to create, but not a validation, i.e.,~checking that the code implements the right algorithm to solve the specific research problem.
Nevertheless, simply attempting to reproduce an output may highlight a submisison's shortcomings in meeting a journal's requirements (cf.~\cite{christian_journal_2020}) and may effectively increase transparency, thereby improving practices (cf.~\cite{nosek_scientific_2012}) even if the check does not go into every detail.

\textbf{Step 4:} The CODECHECKER writes a certificate stating how the code was
run and includes a copy of outputs (e.g.,~data, figures, or tables) that were
independently generated.
The certificate may include further recommendations to the author on how to improve the material.
The free text in the certificate is the most flexible and feasible way to 
make clear what was really checked, because each workflow is unique.
Since no specific tool or platform should be required, such that no authors are excluded, it would be futile for the CODECHECKER to use automation or fixed checklists.

\textbf{Step 5:} The certificate as well as auxiliary files created
during the check, e.g.,~a specification of a computing environment, data 
subsets or helper scripts, and the original code and data get deposited in
an open archive unless restrictions regarding data size, license or 
sensitivity apply.
Currently, the CODECHECKERs deposit the material on Zenodo themselves, but the PUBLISHER may complete this step after integrating CODECHECK into its review process.
A badge or other visual aid may be added to the deposit and the paper and
link to the certificate.
Although a badge simplifies the CODECHECK into a binary value and risks introducing confusion regarding the extent of the check, especially across disciplines and methods, a badge can provide recognition value and acknowledgement of the completed CODECHECK.
The badge and the actual check are incentives for undertaking the effort needed to provide a readily computationally reproducible workflow.

\textbf{Step 6:} The PUBLISHER can, depending on the timing, provide the
certificate to scientific reviewers or handling editors or publish it
and link between the certificate, the paper, and any repositories 
through metadata. Currently, the CODECHECKER creates these connections on
Zenodo\footnote{
See ``related identifiers'' and ``alternate identifiers'' in the right-hand
sidebar of certificate \texttt{2020-025} \cite{cert-2020-025}.}.
The PUBLISHER also gives credit to the CODECHECKER for their work by depositing
the activity appropriately in scholarly profiles, such as ORCID\footnote{
See peer review contributions in ORCID records at
\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}.}.
The PUBLISHER also ensures proper publication metadata, e.g.,~links from the 
certificate repository to the published paper or the original code repository.

@@

The example workflow leaves room for this process to be implemented differently.
To elaborate on these options, we consider several dimensions in a space of possible CODECHECK workflows, as shown in Figure~\ref{fig:dimensions}, to represent relevant aspects for the involved stakeholders.
These aspects touch on timing, responsibilities, and 
transparency and will be discussed in the following sections.
The options on each dimension have their own pros and cons and can,
for the most part, be freely combined.

@@

The point in time at which a CODECHECK is done and its ascribed importance are closely connected, so we describe the dimensions \emph{When} and \emph{Importance} together.
The earlier a CODECHECK happens in the publishing process, the more it can affect editorial decisions:
Is a paper published, sent back for revisions, or rejected?
Even earlier checks, i.e.,~a CODECHECK of a preprint, may even help to improve the workflow itself, even before a publisher is involved.
As such, codechecking papers could be part of a preprint server's policy or simply be initiated by interested readers and community members.

Publishers could introduce a CODECHECK as a \textbf{strict prerequisite}.
As this can reduce the workload of reviewers, such a check should not occur late in the review process.
Yet, the later in the review process the check happens, the easier is it to allow bidirectional communication between the author and codechecker, e.g., because the author might already be notified of the paper's acceptance and may more freely share materials online closer to the paper's publication date.
A \textbf{pre-review} CODECHECK would allow editors to only send a submission out for peer review if it passes the check, or to include the certificate in the submission package provided to the reviewers.
Reviewers may then judge the the relevance of the computations for the results of the work, or follow journal guidelines regarding how strictly to apply the check's outcomes in their evaluation.

A CODECHECK may also be conducted in \textbf{parallel} to the scientific peer review.
This puts less burden on the turnaround time for the check, yet it only makes the outcomes available during the final consideration by the handling editor.
The check could also be assigned after suggestion by a reviewer, which would remove the need for submissions to undergo a pre-review screening.
However, soliciting such a ``specialist review'' is much less desirable than having a regular CODECHECK, thus avoiding the situation in which some submissions get special treatment.
In both cases, a CODECHECK could then be awarded \textbf{equal weight} compared to the other reviews for the editor's decision.

A \textbf{post-acceptance} CODECHECK would have the smallest impact on editorial decisions and may simply provide \textbf{extra merit} on top of the submission's acceptance.
This is the least impactful solution in which all material is still evaluated and the results of the check are properly acknowledged, because the check can be completed before the publication of the paper.
The GIScience group of checks (see below) falls into this category:
By displaying a badge on the volume and article landing pages, the AGILE conference highlights articles whose reproducibility was successfully reviewed.
Similarly, in collaborations with journals, some GIScience have been checked while the authors worked on revisions.

A CODECHECK may also be conducted \textbf{post-publication}, though it would require someone to update the article and article metadata to reference the check; otherwise, readers would presumably not be aware of it.
In general, publishers hesite to make such revisions to published articles.
While this option has the least impact on current publishing practices and enables open discourse, it also has the smallest effect on the original submission and does not acknowledge the importance of reproducible workflows for ensuring good scientific practice.

In general, the timing of a CODECHECK needs to be connected with 
established or typical scholarly communication process.
While more radical ideas do exist for changing the paradigms of how researchers can more granually share their work much
\footnote{For example Octopus, \url{https://science-octopus.org/about}, or Hypergraph, \url{https://www.libscie.org/hypergraph}.},
% mimosa is another platform idea, but no real prototype AFAICT:
% https://projects.invisionapp.com/share/E9Z3RKE7W3P#/screens/436978924
% https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
an evolutionary approach has value, and we do not yet suggest to completely eliminating research papers as a way of sharing work.
Enhancing existing processes with CODECHECKs allows communities to transition towards more open practices at their own pace and ensure that no one is unfaily left behind.
When integrating a CODECHECK into existing review and publication processes, the \emph{turnaround time} is crucial.
Depending on when and who conducts the check, it might be done quickly or it might add considerable time to how long it takes from submission to publication.
In our first checks, we found that a CODECHECK generally takes from 2-5 hours, with some outliers on the higher end.
This time includes writing and publishing the certificate but excludes the time for actual computations.
These efforts are comparable to the time needed to peer review a submission,
which aligns with the efforts some volunteer codecheckers are willing to
make.
However, the time includes a considerable amount of communicating about the
process, especially regarding who publishes which document when, so that proper
cross-referencing between paper and certificate is ensured via persistent
identifiers.
When integrated into a peer review platform, this handling of documents should
become much more effective and reduce organisational overhead.

\subsubsection*{Openness, or ``Who knows who?''}\label{who-knows-who}

The topic of anonymity is broadly discussed, especially in the push towards
open peer review as part of the Open Science movement 
(cf. \cite{ross-hellauer_guidelines_2019}).
Without taking a strong stance on this topic, the general motivation 
behind CODECHECK
for higher transparency and reproducibility does indeed favour a more open 
review process.
However, anonymity can also serve to protect individuals \cite{tennant_limitations_2020}, e.g.,~junior scientists, whereas revealing of names from tenured researchers might be more reasonable to require.
The possible negative effects of a non-anonymous review are reduced if a CODECHECK is not relevant for a journal's decision to accept or reject, but that is, of course, not desirable when the goal is higher transparency and reproducibility.
Instead, we argue that a CODECHECK is technical process that should in general find fixable problems; it is not aimed at giving an opinion or identifying a faulty approach.
It is technically possible for a nonsense workflow to receive
a CODECHECK certificate.
If passing a CODECHECK becomes mandatory, the full transparency may have to be 
revisited as the relations between authors and codecheckers would fall under the
same social and community challenges as non-anonymous peer review
(cf.~\cite{everythinghertz123}).
% previous sentenc inspired by https://osf.io/9cftx/

The technical nature of the check and the challenge of providing sufficient documentation is why we see great benefits in allowing for bidirectional communication between author and codechecker.
Instead of trying to
fix problems or guess the next step, the codechecker can ask the author to 
rework the documentation or the update code.
Instead of struggling to provide perfect instructions and as a result possibly not sharing any code or data, the author can make a best effort to document sufficiently.
Of course, the codechecker may also provide a fix within the code instead
of lengthily describing a small change if it can be quickly done.
Authors and readers can profit from a codecheckers' experience and approach, as during the check they may create useful and instructive files, e.g.,~a machine-readable computing environment specification.
While communication between author and codechecker may be facilitated in an 
anonymous way by the 
publisher, it most likely only helps to protect the identity of the codechecker, because, especially when codebases are of a certain size, not trivial, or if very specialised, they are almost impossible to fully anonymise.
Therefore, the most effective and desirable situation for the stakeholders is to hold a non-anonymous and collaborative CODECHECK.
The contributions by the codechecker may even be integrated into
the code of the workflow and be acknowledged as code commits. This way, 
proper credit can be given within the research software development community.
Nevertheless, depending on the needs represented by other dimensions,
the aspect of ``Who knows who?'' may be the most flexible.
